---
title: "Time_series forecasting on Australian monthly gas production"
output:
  html_document: default
  pdf_document: default
---

Here we are going to predict the gas production for next 12 periods.

# Libraries
```{r}
library(tseries)
library(forecast)
library(ggplot2)
```

Checking the gas production data
```{r}
data("gas")
gas
```

```{r}

autoplot(gas) + labs(x = "Year", y = "Gas Production", title = "Australian monthly gas production from 1956 to 1995")
```
The data is not stationary. It has Level, Seasonal and Trend parameters in it.

```{r}
boxplot(gas ~ cycle(gas), xlab = "Year", ylab = "Gas Production", main = "Australian monthly gas production from 1956 to 1995", col = "gold")
```
From Boxplot it looks like the production of gas is higher in July month.

```{r}
seasonplot(gas)
```



# Making the data stationary
```{r}
plot(diff(log(gas)))
abline(h = 0)
kpss.test(diff(log(gas)))
```

Since the `p-value is 0.1` the data is stationary with first difference. Hence `d = 1`

# Determing p and q values for ARIMA model
```{r}
pacf(diff(log(gas))) # p = 9
```

```{r}
acf(diff(log(gas))) # q = 2  
```

## ARIMA
Running ARIMA model in full data.

```{r}
arima_gas <- auto.arima(gas)
arima_gas
pred <- predict(arima_gas,n.ahead = 12*1)

pred
```

```{r}
forecast_gas <- forecast(arima_gas, h = 12*1)
plot(forecast_gas)
```

```{r}
accuracy(forecast_gas) # MAPE - 3.9
shapiro.test(arima_gas$residuals)
Box.test(arima_gas$residuals,type = "Ljung-Box")
```

Trying out smoothening model

## 1. SES
```{r}
model_gas_ses <- HoltWinters(gas,beta = FALSE, gamma = FALSE)
plot(model_gas_ses)
model_gas_ses #alpha = 0.99
```

```{r}
SES_forecast <- forecast(model_gas_ses,12)
plot(SES_forecast)
accuracy(SES_forecast) #MAPE - 8.22
```

## 2. Holt Trend
```{r}
model_gas_trend <- HoltWinters(gas,gamma = FALSE)
plot(model_gas_trend)
```

```{r}
Trend_forecast <- forecast(model_gas_trend,12)
plot(Trend_forecast)
accuracy(Trend_forecast) # MAPE - 8.15
```

## 3. Holt Seasonal
```{r}
model_gas_Seasonal <- HoltWinters(gas)
plot(model_gas_Seasonal)
```

```{r}
Seasonal_forecast <- forecast(model_gas_Seasonal,12)
plot(Seasonal_forecast)
accuracy(Seasonal_forecast) #MAPE - 4.29
```

Since the `alpha is 0.99` we are taking recent data for ARIMA model
Choosing data from 1991 to 1995 

```{r}
gas_recent <- ts(gas[421:476],start = c(1991,1) , end = c(1995,8), frequency = 12)
plot(diff(log(gas_recent)))
kpss.test(diff(gas_recent))
```

```{r}
pacf(diff(log(gas_recent))) #1
acf(diff(log(gas_recent))) #2
#  d = 1
```

## ARIMA model on 1991 to 1995 year data
```{r}
model_gasrecent_arima <- arima(log(gas_recent),c(1,1,2),seasonal = list(order = c(1,1,2),period = 12))
# arima_gas_recent <- auto.arima(gas_recent)
model_gasrecent_arima
pred <- predict(model_gasrecent_arima,n.ahead = 12)
pred1 <- (2.718)^pred$pred
pred1
```

```{r}
forecast_gas <- forecast(model_gasrecent_arima, h = 12*1)
plot(forecast_gas)
```

```{r}
accuracy(forecast_gas) # MAPE - 0.24
shapiro.test(model_gasrecent_arima$residuals)
Box.test(model_gasrecent_arima$residuals,type = "Ljung-Box")
```
